---
title: "Plotting with ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries
We import tidyverse, readx and here.
```{r libraries, echo=FALSE}
# install.packages("here")
# install.packages("readxl")
# install.packages("tidyverse")
# install.packages("rlang")

library(rlang)
library(here)
library(tidyverse)
library(readxl)

```

### Read from CSV
CSV stands for "comma separated values". Contrary to the name, the separators don't have to be commas!
``` {r csv, echo=FALSE}
ca_np <- read_csv(here("data", "ca_np.csv"))
```

### Read from Excel sheet
An Excel sheet is a sheet from an excel workbook. Nothing misleading about the name this time.
```{r xlsx, echo=FALSE}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

### Plotting with ggplot
We tell R to plot something using ggplot.
```{r plots}
# ggplot(data = ci_np, aes(x = year, y = visitors)) + geom_line()
# # Generally: ggplot(data = my_input, aes(x = x var, y = y var)) + geom_layer() 

ggbasic <- ggplot(data = ci_np, aes(x = year, y = visitors)) # so that we can add geoms on top

# Adding multiple plots in a rmarkdown chunk makes a cool little slideshow for us
# just points
ggbasic + geom_point(aes(size = visitors, color = year), alpha = 0.5) + theme_minimal() + labs(x = "Year", y = "Visitors", title = "Channel Islands NP Visitation", subtitle = "1963 to 2016") + geom_line(color="red")
# columns
ggbasic + geom_col(aes(fill = visitors)) + theme_dark()
# filled shape, has no AA on preview, renders very jagged. should look into this later
ggbasic + geom_area() + theme_linedraw()


```

### Plotting for multiple series
This time we look at *all* parks in California.
```{r allparks}
ggca <- ggplot(data = ca_np,aes(x = year, y = visitors)) + geom_line(linetype = "dashed")
```
