---
title: "Lobsters"
output:
  html_document: default
  pdf_document: default
---
### Installing packages
```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
# attach libraries
library(tidyverse)
library(readxl)
library(here)
library(skimr)  
library(kableExtra)  
```

### Read lobsterdata
Don't forget to ommit the first couple of rows.
```{r read}
lobsters <- read_xlsx(here("data","lobsters2.xlsx"), skip = 4) # skip the first 4 rows
head(lobsters) #see data
# skimr::skim(lobsters) #skim using skimr - packagename::function(argument) <- for readability
head(lobsters) # head of table
```

###Manipulating
Manipulating the tables using dplyr.
```{r dplyr}
lobsters %>% # take our data %then%
  group_by(year) %>% # group it by year %then%
  summarise(count_by_year = n()) #summarise it by yearly count
```
Nice. We can now see how many lobsters were caught *each year*.

###Grouping multiple variables
Manipulating the tables using dplyr - now with multiple groups.
```{r dplyragain}
lobsters %>% # take our data %then%
  group_by(site, year) %>% # group it by site and then year %then%
  summarise(count_by_year = n()) #summarise it by yearly count
```
This answers the question - **how many lobsters were caught in each site in each year?**

### Summarising multiple variables
Let's check the mean value.
```{r meansum}
lobsters %>% # take our data %then%
  group_by(site, year) %>% # group it by site and then year %then%
  summarise(count_by_year = n(),#summarise it by yearly count
            mean_lobster_size = mean(size_mm)) #calculate mean
```
We know the answer to the following question: **what is the average size (in mm) of a lobster caught in each lake in each year?**. Or we would, if not for those pesky missing numbers, where R can't calculate the value of the mean.

### Summarising multiple variables, but better.
Let's check the mean value.
```{r meansumrm}
siteandyear_summary <-
  lobsters %>% # take our data %then% (we also add it to a variable)
  group_by(site, year) %>% # group it by site and then year %then%
  summarise(
  count_by_year = n(),
  #summarise it by yearly count
  mean_lobster_size = mean(size_mm, na.rm = TRUE),
  #calculate mean, remove non-nums
  sd_lobster_size = sd(size_mm, na.rm = TRUE)
  )  #calculate sd, remove non-nums
  
  siteandyear_summary #call
  
  ggplot(data = siteandyear_summary, aes(x = year, y = mean_lobster_size, color = site)) +
  geom_line() #plot
  
  siteandyear_summary %>% kable() #put into table for markdown
```
In total there are `r nrow(lobsters)` included in our report.

### Mutating and selecting.
Just adding columns. The table "grows" with specified instructions. Select selects.
```{r table}
lobsters_mutated <- lobsters %>% mutate(size_cm = size_mm / 10, # size in centimeters
                    caught_by = "Lobsterfisher") # who caught it? 

lobsters_mutated %>% select(date, site, size_cm)

```

### Filtering and joining
This is all chapter eight, but in the same .rmd.
Filtering is almost like VLOOKUP from Excel (filters by rows)
```{r filterjoin}
fish <- read.csv(here("data","fish.csv"))
kelp_abur <- read_excel(here("data", "kelp.xlsx"),sheet="abur")

# calls

kelp_abur

fish_garibaldismith <- fish %>% filter(str_detect(common_name, pattern = "black"))
fish_garibaldismith


abur_kelp_fish <- kelp_abur %>% inner_join(fish, by = c("year","site"))
abur_kelp_fish

```

